---
title: "The Search for the Best NFL Defenses (1978-2021)"
author: "Oscar Monroy"
date: "3/20/2022"
output: 
  html_document:
    theme: darkly
    df_print: paged
---

# Table of Every NFL Team's Defensive Stats from 1978 to 2021

```{r, message=FALSE}
library(dplyr)
def <- read.csv("nfl_def.csv")
names(def)[1] <- "Team"
def
```

# Finding the Greaest Defenses

```
Generally, people consider the '85 Chicago Bears to be the best defense of all time. 
Without a doubt they were one of the best to ever do it, but were they 
truly number 1? The 1985 Chicago Bears played in an era where passing plays were considered 
a novelty and running backs were the main players that moved the ball forward, so they 
specialized in creating immense pressure in the backfield and terrorizing running backs. They
weren't too shabby at passing defense, but that wasn't the main point of that defense since 
they didn't need to prepare much for that (except against Dan Marino's Miami Dolphins, who 
the Bears lost to). Nowadays, if you were to play the '85 Bears in today's game, they would 
more than likely get absolutely dominated as the game has evolved more in the direction that 
Dan Marino had played in: pass happy. Nevertheless that Bears team is still considered 
one of the greatest for a good reason and part of that is because they hardly allowed their
opponents to score. However there are more teams that people overlook when talking about 
NFL defenses, and we're about to see who they are with some simple data analysis.
```
# Teams that Just Didn't Like Points Being Scored on Them

```{r}
# First let's grab the data for total points allowed for teams that
# ranked first in their respective year.
num1 <- def %>%
  filter(Rank %in% c(1)) %>%
  group_by(Team)
num1
# We're certainly seeing an abundance of Steelers, 49ers, and Vikings on this list...

# Let's get a look at who the top dogs are in total 
# points allowed in their respective year.
num1 %>%
  select(Team, Year, Gms, TotPts) %>%
  arrange(TotPts)
# Now here is an interesting table with an unexpected #1 defense: the '82 Dolphins.
# Now I put in the the Gms variable (Games Played) in here to show an obvious disparity.
# During the 1982, the Players' Strike was going which dropped the number of games
# that season, so ultimately, we end up with this strange bit of info where the
# Miami Dolphins appeared to have achieved one of the greatest feats in sports
# history at first glance until you realize there was only 9 games.Aside from that,
# we still see the '85 Bears in the Top 5 (Top 4 if your exclude the Dolphins),

# Here's is another interesting factoid: the '86 Bears were actually better in every way
# than the '85 Bears. However, due to a malicious injury caused to QB Jim Mcmahon, their
# season was completely derailed and now the '86 counterpart is mostly forgotten. If that
# injury didn't happen, they would be remembered far more than they currently are.

# Also, the Tennessee Titans (my favorite team) comes in at 3rd place (or 2nd because...
# you know) and also could've been considered one of the greatest teams in history if 
# they weren't stopped in the playoffs by rival Baltimore Ravens.

# Just for fun, let's see the '82 Dolphins stats:
num1[40, ]
# Using their Points Allowed per Game stat, we can estimate where their TotPts would've been
# had they played 7 more games for a standard 16 game season:
131 + (14.6 * 7)
# Not too shabby. At 233.2, they would've ranked at #15 in the list
# of best defensive teams of their respective year.

# You may have notcied that I was using the limitation of only having the
# number 1 rated defense of their year but that alone doesn't tell a story. A team could
# be experts at preventing touchdowns from happening but they could also
# simultaneously "struggle" at preventing the ball from traveling to their side
# of the field...
t <- def %>%
  filter(Year != 1982) %>%
  arrange(TotPts)
head(t, 10)
# and the team that fits that bill is the 2000 Baltimore Ravens. They were experts 
# at preventing points from being scored but still allowed the opposition to gain yards.
# It is likely that they were just exceptional at the red-zone defense and/or 
# getting forced turnovers. Regardless, you'll notice how they're polar opposites
# of their rival from that same year: the Tennessee Titans. Titans allowed more points
# but allowed the least yards [not to say they were bad (far from it) at preventing 
# points, but it really is interesting at the opposite method both teams' defenses had].
# Another anomaly is the 1978 Denver Broncos who rank 7th overall and 6th best defense
# of that particular year. Just look at the total yards they allowed; it's like 
# an extreme version of the 2000 Ravens. What even was happening there?
# Oh and the legendary Steel Curtain Steelers rank #4 overall but #3 of that year.
# Honestly, what was going on in 1978 and 2000? Maybe we'll find out.
```

## Winner of the "Team that Didn't Like Points Being Scored on Them" Award:
## The 2000 Baltimore Ravens
### Not bad for a team that ranked #2 in that same year. I still don't like them.

# Teams that Weren't Fans of Seeing their Opponents Move

```{r}
# I'm going back to using the rank 1 defenses of the year as the limitation
# again only because I like surprising people when I do remove the restriction.
num1 %>%
  select(Team, Year, Gms, TotYds) %>%
  filter(Year != 1982) %>%
  arrange(TotYds)
# First of all, the '85 Bears rank #19 and I blame Dan Marino for that.
# Secondly, we see a load of fresh new faces on the top 10. The '91
# Philadelphia Eagles had a monstrous defense that forced their rivals
# New York Giants, Washington Redskins, and Dallas Cowboys
# to use elite offensive linemen and while the defense still held up
# incredibly well, they still fell short of the Super Bowl. To add
# insult to injury, their rival from New York ultimately won the big game.
# (Granted the New York Giants had help from the greatest defensive 
# player to ever exist: Lawrence Taylor, but that's a story for another time)

# Also, look at that. We finally know who #1 in 1978 was: the LA Rams.

# Time to remove the restrictions
t2 <- def %>%
  filter(Year != 1982) %>%
  arrange(TotYds)
head(t2, 10)
# Turns out there weren't any surprises this time. The top 10
# is expectedly filled with rank 1 defenses except for the
# 1991 New Orleans Saints, 2000 Baltimore Ravens and, 
# 1978 Dallas Cowboys. This only reinforces that 1978 was the
# wildest year for NFL defenses and also maybe 2000. The 2008
# Pittsburgh Steelers being here is also neat considering that
# defense really came in handy for that Super Bowl win.
```


## Winner of the "Team that Didn't Take too Kindly to Seeing Movement" Award:
## The 1991 Philadelphia Eagles
### At least they won their Super Bowl 26 years later.

```
Well we've seen the best historical defenses when we account for the most
generalized stats like Total Points Allowed in a Season and Total Yards
Allowed in a Season. Of course, the teams that came out on top when only looking 
at these stats are truly the greatest that we'll ever see. But generalized stats
alone aren't fun; I want to look at teams that specialized in specific defensive
aspects of the game like the rushing defense and passing defense. I'd like to make
some predictions before we get into that though. Prediction 1: the best rushing defense
will consist of those incredible 1978 defenses that defined the year and also
the '85 Bears. Prediction 2: the best passing defenses will consist of 21st century
era NFL defenses and contain 2013-2014 Seattle Seahawks, any of the Ed Reed era
Baltimore Ravens, and the 2015 Denver Broncos. 
```

# Teams that Just Had to Stuff the Run

```{r}
# Of course, we'll start out with the run defense as the run was the
# bread and butter of 20th century (and 21st century Tennessee Titans)
# offenses. Passing offenses were scarce in the 20th century, so we'll 
# leave that for last. We'll also do the number 1 defenses of their year first
# just for that surprise factor when people realize that the best defense
# doesn't actually have the best rushing defense, maybe.
num1 %>%
  select(Team, Year, Gms, RushYds) %>%
  filter(Year != 1982) %>%
  arrange(RushYds)
# Hey look! The '85 Bears are at 10th place.
# Anyways, you'll notice that (in the top 10 at least), that there is an
# even split of 20th century and 21st century era defenses. And the team
# with the best rushing defense overall is the '95 San Francisco 49ers, who
# fielded and incredible run stuffing defense that barely allowed over 1000
# yards and an offense led by the main QB himself Steve Young. We also see
# the legendary '91 Philadelphia Eagles and a surprising new face: the 
# '98 San Diego Chargers, who lost 3 years ago in the Super Bowl to our number 1
# overall rushing defense. The surprising 21st century era rushing defenses consist
# the eternal AFC North rivals: the Pittsburgh Steelers and Baltimore Ravens, both 
# of whom were in a "arms race" of sorts to create the best defense to beat one
# another with. Then there is the 2014 Legion of Boom Seattle Seahawks, who had 
# such a ferocious defense in general that they earned the cheesy, yet intimidating
# nicknane Legion of Boom. This was also the team that tragically lost to the 
# New England Patriots on a last second interception because they didn't run the ball.

# Anyways, let's remove the rank-1-only restrictions.
t3 <- def %>%
  filter(Year != 1982) %>%
  arrange(RushYds)
head(t3, 10)
# This whole list is surprisng. First of all, I would've never expected to see
# the Detroit Lions in a top 10 list of the best of anything (unless it's
# a list of the top 10 saddest fanbases). The 2000 Baltimore Raven once again
# reaches the top of the list, only allowing a measley 970 yards. Then we have 
# the Minnesota Vikings who ranked number 8 best defense of their own year yet
# fielded the second best rushing defense ever, only allowing 985 yards. We have yet 
# another new face in the 2010 Steelers before we finally reach the '95 SF 49ers.
# Well, if you wanna know how the 2006 Vikings made number 2 overall despite being
# number 8 for their year, just look at their rushing yards allowed stats and then look 
# at their passing yards allowed and total points allowed stats. Absolutely
# beqildering, yet in character for the perpetually cursed Minnesota franchise.
# Not as bewildering as the Lions being here. That is still a "What the hell? moment.
```

## Winner of the "Best Run Stuffing Team" Award:
## The 2000 Baltimore Ravens
### ...bastards.

## Honorary award goes to the 2014 Detroit Lions!
### It can't be easy to root for the Lions but at least here's this award. 

# The Anti Air Teams

```{r}
# We're now gonna look at the best pass defenses in NFL history and
# we'll follow the same formula we always do.
num1 %>%
  select(Team, Year, Gms, PassYds) %>%
  filter(Year != 1982) %>%
  arrange(PassYds)
# I don't know if it's obvious, but there is a problem here.
# Passing offenses didn't become popular until later on and
# it was a slow process. Just look at the increasing amount of
# pass attempts done per year in one of my previous bodies of
# works (about NFL passing offenses). According to New York Tines,
# "In 1980, teams pass the football 47 percent of the time, compared 
# with just 38 percent in 1977...". So my main point here is that there
# is gonna be some skewed numbers looking at older team because they
# didn't have to deal with much passing offenses back then.

# The only way (I believe) to truly look at who the best passing defense is to
# have a "passes allowed" to "total passes done against the team" stat.
# Unfortunately, we don't have such a stat at the moment. Best I can
# do is maybe set the Year variable to start from 1984 (when Dan Marino
# started popularizing the passing offense) and also just brag that the
# 2000 Tennessee Titans are still on the top 10 list despite the issue.
num1 %>%
  select(Team, Year, Gms, PassYds) %>%
  filter(Year > 1983) %>%
  arrange(PassYds)
# If we look at the top 3, we see that the 1991 Philadephia Eagles once again
# reigns supreme with the 2000 Tennessee Titans coming in second. Now for the
# big surprise, we get the New York Jets coming in at number 3 which is 
# absolutely crazy considering they had to deal with Tom Brady twice a year.

# But now we remove the ranking restrictions.
t4 <- def %>%
  filter(Year > 1983) %>%
  arrange(PassYds)
head(t4, 10)
# And now we have a new number 1 overall passing defense: the 2006 Oaklands 
# Raiders. The Raiders haven't been a steady franchise ever since they lost the
# Super Bowl in 2002, so to see them after 2002 here is quite a shock. We do
# see some shifts in rankings such as the addition of '88 Kansas Chiefs (who
# ranked number 10 in 1988) and the '84 New Orleans Saints. Every team ranked
# after the '88 Vikings in the previous top 10 list got push out of the top 10
# list here without the restrictions. 
```

## Winner of the "Anti Air Team" Award:
## The 2006 Oakland Raiders
### Maybe you'll win one soon... in Las Vegas. Sorry Oakland.
